---
title: "Assignment04"
author: "Alexa Beeson acb343"
format: pdf
---

# Data Source: American Community Survey via [IPUMS USA](https://usa.ipums.org/usa/)

Loading in relevant programs

```{r}
library(ipumsr)
library(tidyverse)
library(haven)
library(ggplot2)
library(srvyr)
```

Reading in the data set:

```{r}
setwd("data")
ddi <- read_ipums_ddi("usa_00002.xml")
data <- read_ipums_micro(ddi)
```

Applying the appropriate weights to the sample:

```{r}
wdata <- data %>%
    as_survey_design(weights=HHWT)
```

# Data Visualization 1:

Counting the number of insured adult respondents
```{r}
data %>%
  filter(AGE>18) %>%
    mutate(insurance=
           case_when(
             HCOVANY==1 ~ "Uninsured",
             HCOVANY==2 ~ "Insured"
           )) %>%
  group_by(insurance) %>%
  count(insurance)
```
Creating a plot to summarize the results:
```{r}
options(scipen=999)
data %>%
  filter(AGE>18) %>%
  mutate(insurance=
           case_when(
             HCOVANY==1 ~ "Uninsured",
             HCOVANY==2 ~ "Insured"
           )) %>%
  ggplot(aes(x=insurance))+
  geom_bar(stat="count") +
  labs(
    x="Insurance Status",
    y="Observations of Insurance Categories",
    title="A Count of the Number of Adults with and without Insurance",
    subtitle="Comparing the number of insured to uninsured respondents in 2019"
  )+
  scale_x_discrete()+
  theme_light()
```
# Visualization Description:
This first data visualization helps us break down the population of respondents by whether or not they are insured. This bar graph depicts the count of people who are insured and compares it to the number of people who are uninsured. We can see that while over 2.3 million respondents are insured, there are still over 200,000 who are not. If we just examine the raw count of respondents with health insurance, it may trick us into underestimating the true number of people who remain uninsured despite changes to our health insurance systems, like expanding Medicaid, offering subsidies to buy insurance, and creating stronger regulations on private insurance. Insurance status in the United States can literally mean life or death for a sick person, so it is important to dive further into those 220,000+ uninsured respondents to see if there are any patterns in people who remain uninsured.